NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-04-20 Manuscript/manuscriptPS.tex
update manuscript

\caption{List of MD simulations. The salt concentrations calculated as 

@@

    POPS/Na$^+$  & MacRog \cite{maciejewski14}  &0 & 0        & 128 & 5120 & 0  & 298  & 200 & 100 & \cite{macrogPOPS298K} \\

@@

@@

@@

The headgroup order parameters of DOPS and POPS bilayers 
from different simulation models are compared with the experimental
data in Fig. \ref{HGorderParametersPS}. In line with the previous
study for PC lipids \cite{botan15}, the glycerol backbone order 
parameters in CHARMM36 roughly agree with the experimental data,
while significant discrepancies for glycerol backbone carbons
are observed in other models.
\todo{Discussion to be finished once all the results are in the plot.} \\
\todo{Also the discussion about POPS/OPPS issue with MacRog model should be added.}

Based on the subjective ranking shown in Fig. \ref{comparisonTablePS},
the CHARMM36 is the best performing model for both lipids.
However, none of the tested models give a satisfactory 
agreement with experiments for the headgroup order paramaters of 
PS lipids. The total deviation from the experiments ($\Sigma$ in Fig. \ref{comparisonTablePS})
for the best performing CHARMM36 model is larger for PS lipids (8)
than for PC (3) \cite{botan15}. Therefore the 
interpretation of structural details of PS headgroup or
differences between PC and PS lipids is very challenging 
from the current MD simulation models.
Figures \ref{dihedralsGLY} and \ref{dihedralsHG} show dihedral
angle distributions calculated from different models. 
The glycerol backbone structures from CHARMM36 and
Slipids simulations visualized in Fig. \ref{glycerol_buslaev} reveal the
missing structures in Slipids model, which probably lead to the incorrect
order parameters. For the headgroup $\alpha$ and $\beta$ carbon order parameters of PS lipids,
the tested models perform less well than for PC headgroup in previous
study \cite{botan15}.
\todo{Discussion will be finished when we have all the data in
  Figs. \ref{HGorderParametersPS}, \ref{dihedralsGLY} and \ref{dihedralsHG}.
  One possible conclusion could be the following:
  The main differences between the models in the headgroup region are observed
  for dihedrals C12-C11-O12-P and C11-C12-C13-O1A.
  CHARMM36, CHARMM36UA and Slipids give very similar results to the
  dihedral C11-C12-C13-O1A, which is close to the $\beta$-carbon.
  The order parameters of $\beta$-carbons for these three models
  are in best argeement with the experiments in figure \ref{HGorderParametersPS}.
  On the other hand, Gromos-CKP models give better order parameters
  for $\alpha$-carbon than Slipids, CHARMM36 or CHARMM36UA. 
  In conclusion, the suggestion would be that the single peak for
  observed at 120 degrees in CHARMMs and Slipids would
  be more realistic for C11-C12-C13-O1A dihedral, while the single peak at 180 degrees observed
  in CKP models and in Berger would be most realistic for C12-C11-O12-P dihedral.
}

@@

\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/CIdensPSOCmixt.eps}
  \caption{  Counterion density distributions from PC:PS mixtures.
\label{CIdensPSOCmixt}
  }
\end{figure}

@@

>>>>

\caption{List of MD simulations without additional salt.

@@

    POPS/Na$^+$  & MacRog \cite{maciejewski14}  &0 & 0        & ? & ?? & 0  & ?  & ? & ?  & \cite{??} \todoi{Data to be added by Piggot} \\
    OPPS/Na$^+$  & MacRog \cite{maciejewski14}  &0 & 0        & 128 & 5120 & 0  & 298  & 200 & 100 & \cite{macrogPOPS298K} \\

@@

\caption{  List of MD simulations. The salt concentrations calculated as 
   'M' stands for Mukhopadhyay after the first published Berger-based PS simulation that used these charges \cite{??})

@@

    \todo{MacRog simulations with KCl to be added}\\
    \todo{Berger simulations with NaCl and CaCl to be added}

@@

The headgroup order parameters of DOPS and POPS bilayers 
from different simulation models are compared with the experimental
data in Fig. \ref{HGorderParametersPS}. In line with the previous
study for PC lipids \cite{botan15}, the glycerol backbone order 
parameters in CHARMM36 roughly agree with the experimental data,
while significant discrepancies for glycerol backbone carbons
are observed in other models.
\todo{Discussion is to be finished.} \\
\todo{Also the discussion about POPS/OPPS issue with MacRog model should be added.}

Based on the subjective ranking shown in Fig. \ref{comparisonTablePS},
the CHARMM36 is the best performing model for both lipids.
However, none of the tested models give a satisfactory 
agreement with experiments for the headgroup order paramaters of 
PS lipids. The total deviation from the experiments ($\Sigma$ in Fig. \ref{comparisonTablePS})
for the best performing CHARMM36 model is larger for PS lipids (8)
than for PC (3) \cite{botan15}. Therefore the 
interpretation of structural details of PS headgroup or
differences between PC and PS lipids is very challenging 
from the current MD simulation models.
Figures \ref{dihedralsGLY} and \ref{dihedralsHG} show dihedral
angle distributions calculated from different models. 
The glycerol backbone structures from CHARMM36 and
Slipids simulations visualized in Fig. \ref{glycerol_buslaev} reveal the
missing structures in Slipids model, which probably lead to the incorrect
order parameters. For the headgroup $\alpha$ and $\beta$ carbon order parameters of PS lipids,
the tested models perform less well than for PC headgroup in previous
study \cite{botan15}.
\todo{Discussion is to be finished.
  One possible conclusion could be the following:
  The main differences between the models in the headgroup region are observed
  for dihedrals C12-C11-O12-P and C11-C12-C13-O1A.
  CHARMM36, CHARMM36UA and Slipids give very similar results to the
  dihedral C11-C12-C13-O1A, which is close to the $\beta$-carbon.
  The order parameters of $\beta$-carbons for these three models
  are in best argeement with the experiments in figure \ref{HGorderParametersPS}.
  On the other hand, Gromos-CKP models give better order parameters
  for $\alpha$-carbon than Slipids, CHARMM36 or CHARMM36UA. 
  In conclusion, the suggestion would be that the single peak for
  observed at 120 degrees in CHARMMs and Slipids would
  be more realistic for C11-C12-C13-O1A dihedral, while the single peak at 180 degrees observed
  in CKP models and in Berger would be most realistic for C12-C11-O12-P dihedral.
}

@@

\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/CIdensPSOCmixt.eps}
  \caption{  Counterion density distributions from PC:PS mixtures.
\label{CIdensPSOCmixt}
  }
  \todo{Fix CKP1 and CKP2 captions.}
\end{figure}

@@

\subsection{CHARMM36}
\todo{To be written by Piggot, Madsen and Ollila}

\subsection{CHARMM36ua}
\todo{To be written by Piggot}

\subsection{Slipids}
\todo{To be written by Piggot and Favela}

\subsection{Berger}
\todo{To be wiritten by Piggot and Ollila}
Simulations with sodium were taken directly from Ref. \citenum{??} and
simulations with calcium directly from \citenum{melcrova16}.

\subsection{GROMOS-CKP}
\todo{To be written by Piggot}

\subsection{Lipid17}
\todo{To be written by Kav and Miettinen}

\subsection{MacRog}
\todo{To be written by Javanainen and Piggot}
